Functory: Distributed Computing for the Common Man\thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2010-10-07 doc/report/main.tex
*** empty log message ***

\section{Illustrative Case Studies}

In this section, we show how to use \functory\ on several case studies.
We focus here on the use of the API. Experimental results of these
case studies are given in Section~\ref{sec:experiments}.

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to p-1 do
      for k = 0 to m-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, assuming constant
time addition and multiplication (denoted $+$ and $\times$).
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to p-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

@@

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string t = string_of_int (worker (task_of_string i))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode integers
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- int_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

@@

@@

Our library, \functory, is implemented in \Ocaml\ and is available
from \url{http://www.lri.fr/~filliatr/functory/}.
In this section, we show the potential of this library using two 
computationally extensive examples.

\subsubsection{Matrix Multiplication}

$\of{n} = \of{p} = \of{m} = 100$

coefficients have several thousands of digits; we use GMP

%            parall  line

% sequential  20.3   20.2
% cores 2     22.7   11.3
% cores 4     12.3    6.1
% cores 6      8.6    4.3
% cores 8      8.0    3.5

% network is localhost(moloch) + belzebuth

% network 3+3        12.7
% network 6+  112     7.2
% network  +6   ?    36.4
% network 6+6  76.1  11.9

\begin{center}
  \begin{tabular}{|l|l|l|}
    \hline
    sequential & cores & network \\\hline
    20.3 & 
  \end{tabular}
\end{center}

@@

% TODO compute ratios for Network
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    \#cores  &\#tasks & \of{Cores} & \of{Network} \\
    \hline
    2       & 10 & 15.8       (1.86) &  20.3       \\
            & 30 & 15.7       (\underline{1.87}) &  18.7       \\
            & 100 & 16.1      (1.83) &  19.8       \\
            & 1000 & 19.6     (1.50) &  38.6      \\
    \hline
    4       & 10 & 9.50       (3.09)  &  14.4       \\
            & 30 & 8.26       (\underline{3.56})  &  11.4   \\
            & 100 & 8.37      (3.51)  &  11.4   \\
            & 1000 & 10.6     (2.77)  &  20.5    \\
    \hline
    8       & 10 & 9.40       (3.13)  &  12.6      \\
            & 30 & 4.24       (\underline{6.93})  &   7.6      \\
            & 100 & 4.38      (6.71)  &   7.5      \\
            & 1000 & 6.86     (4.29)  &  11.3      \\
    \hline
  \end{tabular}
\end{center}
The best timings are achieved for the \of{Cores} configuration, where
communications happen within the same machine and are thus cheaper.
There are two significant differences with respect to the n-queens
benchmark.  On one hand, the number of tasks can be controlled more
easily than in the case of n-queens. We experimentally figured out the
optimal number of tasks to be 30. One the other hand, each computation
result is an image, rather than just an integer as in the case of
n-queens. Consequently, communication costs are much greater. 
In this particular experiment, the total size of the results
transmitted is more than 200 Mb.

@@

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. Our computing
infrastructure consists of 3 machines with 4, 8 and 8 cores
respectively. The benchmark we use for
experimentation consists of around 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
which are all installed on each of the machines in use.

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 1 minute.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{unknown} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r||r|r|r|r|}
    \hline
    prover   & valid & unknown & timeout & failure
    \\\hline\hline
    Alt-ergo & 406.0 & 3.0   &  11400.0 & 0.0       
    \\\hline
    Simplify &  0.5   & 0.4   &  1200.0 & 222.0   
    \\\hline
    Z3       & 80.7   & 0.0   &  1800.0 & 1695.0   
    \\\hline
    CVC3     & 303.0  & 82.7  &  4200.0 & 659.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 6 hours if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 22 minutes and 37 seconds, giving us a
speedup of more than 16$\times$. We are still far away from the ideal
ratio of 20$\times$ (we are using 20 cores), since some provers are
allocating a lot of memory and time spent in system calls is not
accounted for in the total observed time. However, a ratio of
16$\times$ is already a significant improvement for our day-to-day
experiments.

@@

%\vfill\pagebreak
% We recommend abbrvnat bibliography style.
\bibliographystyle{abbrvnat}
\bibliography{./biblio}

>>>>

\section{Illustrative Case Studies}\label{sec:studies}

In this section, we show how to use \functory\ on several case studies.
We focus here on the use of the API. Experimental results of these
case studies are given in Section~\ref{sec:experiments}.
The source code for all these case studies is contained in the
distribution, in sub-directory \texttt{tests}.

\subsection{Matrix Multiplication}\label{sec:matrix}

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to p-1 do
      for k = 0 to m-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
assuming addition and multiplication over coefficients to be 
denoted $+$ and $\times$ respectively.
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to p-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

@@

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string t = string_of_int (worker (task_of_string i))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

@@

\subsection{SMT Solvers}\label{sec:SMT}

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. In this case study,
we consider 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify, Z3 and CVC3),
which are all installed on each of the machines in use.

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file, within
a timeout limit.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{unknown} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.

The result of each computation is a pair denoting the status and the
time spent in the prover call. The master collects these results and
sums up the timings for each prover and each possible status.

@@

Our library, \functory, is implemented in \Ocaml\ and is available
from \url{http://www.lri.fr/~filliatr/functory/}.
In this section, we show the experimental results pertaining to the
case studies of Section~\ref{sec:studies}.

@@

% TODO compute ratios for Network
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    \#cores  &\#tasks & \of{Cores} & \of{Network} \\
    \hline\hline
    2       & 10 & 15.8       (1.86) &  20.3       \\
            & 30 & 15.7       (\underline{1.87}) &  18.7       \\
            & 100 & 16.1      (1.83) &  19.8       \\
            & 1000 & 19.6     (1.50) &  38.6      \\
    \hline
    4       & 10 & 9.50       (3.09)  &  14.4       \\
            & 30 & 8.26       (\underline{3.56})  &  11.4   \\
            & 100 & 8.37      (3.51)  &  11.4   \\
            & 1000 & 10.6     (2.77)  &  20.5    \\
    \hline
    8       & 10 & 9.40       (3.13)  &  12.6      \\
            & 30 & 4.24       (\underline{6.93})  &   7.6      \\
            & 100 & 4.38      (6.71)  &   7.5      \\
            & 1000 & 6.86     (4.29)  &  11.3      \\
    \hline
  \end{tabular}
\end{center}
The best timings are achieved for the \of{Cores} configuration, where
communications happen within the same machine and are thus cheaper.
There are two significant differences with respect to the n-queens
benchmark.  On one hand, the number of tasks can be controlled more
easily than in the case of n-queens. We experimentally figured out the
optimal number of tasks to be 30. One the other hand, each computation
result is an image, rather than just an integer as in the case of
n-queens. Consequently, communication costs are much greater. 
In this particular experiment, the total size of the results
transmitted is more than 200 Mb.

This benchmark was inspired by the PASCO'10 programming contest~\cite{PASCO}.
It consists of multiplication of
two square matrices of dimension 100, that is $\of{n} = \of{p} =
\of{m} = 100$, with integer coefficients.
Coefficients have several thousands of digits, hence we use
GMP~\cite{GMP} to handle operations over coefficients.

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7  &  11.3\\
 (4 cores)     &   12.3  &   6.1\\
 (6 cores)     &    8.6  &   4.3\\
 (8 cores)     &    8.0  &   3.5\\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.

A less naive implementation would have the worker read the input matrices
only once, \emph{e.g.} from a file, and then have the master send only
row and column indices. This would reduce the communication cost to
$O(\of{n}\times\of{m})$ only.

%             mm1     mm2

% sequential  20.3   20.2

% cores 2     22.7   11.3
% cores 4     12.3    6.1
% cores 6      8.6    4.3
% cores 8      8.0    3.5

% network = workers on moloch + master on belzebuth

% cores 2     93     32.3
% cores 4     87     33.7
% cores 6     92     33.5
% cores 8     88     32.5

@@

As explained in Section~\ref{sec:SMT}, this benchmark consists of 80
verification conditions, each being checked by 4 different SMT
solvers. Each task is executed with a timeout limit of 1 minute.
Our computing
infrastructure for this experiment consists of 3 machines with 4, 8 and 8 cores
respectively. 
The figure below shows the total time in minutes spent by each prover
for each possible outcome.
\begin{center}
  \begin{tabular}{|r||r|r|r|r|}
    \hline
    prover   & valid & unknown & timeout & failure
    \\\hline\hline
    Alt-ergo & 406.0 & 3.0   &  11400.0 & 0.0       
    \\\hline
    Simplify &  0.5   & 0.4   &  1200.0 & 222.0   
    \\\hline
    Z3       & 80.7   & 0.0   &  1800.0 & 1695.0   
    \\\hline
    CVC3     & 303.0  & 82.7  &  4200.0 & 659.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 6 hours if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 22 minutes and 37 seconds, giving us a
speedup of more than 16$\times$. We are still far away from the ideal
ratio of 20$\times$ (we are using 20 cores), since some provers are
allocating a lot of memory and time spent in system calls is not
accounted for in the total observed time. However, a ratio of
16$\times$ is already a significant improvement for our day-to-day
experiments.

@@

%\vfill\pagebreak
\bibliographystyle{plain}
\bibliography{./biblio}
